---
title: "Additional examples"
author: "Justin M Shea"
date: ' '
output:
  rmarkdown::html_document:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Additional examples}  
  %\VignetteEngine{knitr::rmarkdown}  
  %\VignetteEncoding{UTF-8}
---


```{r, message = FALSE, warning = FALSE}
library(xts)
library(knitr)
library(neverhpfilter)
```




## yth_filtered

The function outputs an xts object of relevant series discussed in Hamilton (2017).
They are the original, trend, cycle, and random walk series.

```{r, warning=FALSE, message=FALSE}
data(GDPC1)
log_RGDP <- 100*log(GDPC1)

gdp_filtered <- yth_filter(log_RGDP, h = 8, p = 4)

kable(tail(gdp_filtered, 8))
```

As the output is an xts object, it can be easily graphed with `plot.xts`. Note the `panels` argument, which I use to plot the cycle component.

```{r, warning = FALSE}
main <- "Log of Real GDP (GDPC1) and trend"
plot(gdp_filtered["1950/"][,1:2], grid.col = "white", legend.loc = "topleft", main = main,  panels = 'lines(gdp_filtered["1950/"][,3], type="h", on=NA)')
```

While `xts` has many useful functions for handling time series, one can extend the use of this class further with packages such as `PerformanceAnalytics`, Recession bars are created easily with the `period.areas` argument of the `chart.TimeSeries` function.

```{r, warning = FALSE, message = FALSE}
library(PerformanceAnalytics)

  qtr_recc <- to.quarterly(USREC["1947-1/2017-3"], OHLC=FALSE)
     start <- index(qtr_recc[which(diff(qtr_recc$USREC)==1)])
       end <- index(qtr_recc[which(diff(qtr_recc$USREC)==-1)])
Recessions <- paste(format(start,"%Y-%q"), format(end,"%Y-%q"), sep="/")

main <- "Log of Real GDP (GDPC1) and trend"
chart.TimeSeries(gdp_filtered[,1:2], main = main, date.format = "%Y", period.areas = Recessions, period.color="lightgrey", lwd=2)
```



## Choices for h and p

While Hamilton converts the PAYEMS series into quarterly periodicity and then uses
the default parameters of `h = 8` and `p = 4`, one can choose to retain the monthly series and adjust the parameters to `h = 24` and `p = 12`.

```{r, warning = FALSE}
Employment_log <- 100*log(PAYEMS["1950/"])

employment_cycle <- yth_filter(Employment_log, h = 24, p = 12, output = "cycle")

main <- "Log of Employment cycle"
cols <- c("darkred", "darkgreen")[(employment_cycle > 0) + 1]  
plot(employment_cycle, grid.col = "white", type = "h", col = cols, main = main)
abline(h=0)
```

One may wish to consider longer-term cycles by extending `h`. Below is an example of taking h from 8 quarters or 2 years, to 20 quarters (5 years), and then 40 quarters (10 years). 
```{r}
gdp_5yr <- yth_filter(log_RGDP, h = 20, p = 4, output = c("x", "trend", "cycle"))

main <- "Log of Real GDP and 5-year trend "
plot(gdp_5yr["1980/"][,1:2], grid.col = "white", legend.loc = "topleft", main = main,  panels = 'lines(gdp_5yr["1980/"][,3], type="h", on=NA)')

gdp_10yr <- yth_filter(log_RGDP, h = 40, p = 4, output = c("x", "trend", 
"cycle"))

main <- "Log of Real GDP and 10-year trend "
plot(gdp_10yr["1980/"][,1:2], grid.col = "white", legend.loc = "topleft", main = main,  panels = 'lines(gdp_10yr["1980/"][,3], type="h", on=NA)')

```

## yth_glm

The `yth_glm` function returns an object of class `glm`. Thus, one can use all
generic functions associated with objects of that class. Here is an example of
passing the results of a `yth_glm` model to the plot function.


```{r, warning=FALSE, message=FALSE}
gdp_model <- yth_glm(log_RGDP["1960/"], h = 8, p = 4)

plot(gdp_model)
```


Optional parameters supported by the `glm` function can be used as well. An example using a different error distribution, defined with the `family` argument, is illustrated below.

```{r, warning=FALSE, message=FALSE}
gdp_poisson <- yth_glm(log_RGDP["1960/"], h = 8, p = 4, family = poisson)

summary(gdp_poisson)
```

