---
title: "Getting started"
author: "Justin M Shea"
date: ' '
output:
  rmarkdown::html_document:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Additional examples}  
  %\VignetteEngine{knitr::rmarkdown}  
  %\VignetteEncoding{UTF-8}
---


# Introduction

The `neverhpfilter` package consists of 2 `functions`, 12 economic data sets, 
Robert Shiller's U.S. Stock Markets and CAPE Ratio data from 1871 through 2019, 
and a `data.frame` containing the original filter estimates found on table 2 of [Hamilton (2017) <doi:10.3386/w23429>.](https://www.nber.org/papers/w23429) All data objects are
stored as `.Rdata` files in eXtensible Time Series (`xts`) format. 

One of the first things to know about the `neverhpfilter` package is that it's
functions accept and output, `xts` objects. The author uses `xts` often as it 
remains one of the most efficient data objects for managing times series data in 
econometrics and finance. 

An `xts` object is a `list` consisting of a `vector` index of some date/time class
paired with a `matrix` object of singular data type. In our case, type `numeric`.
`data.table` is also heavily used in finance and has efficient date/time indexing 
capabilities as well. It is useful when working with large data.frame like lists 
containing vectors of multiple data types of equal length. If using `data.table` 
or some other index based time series data object, merging the `xts` objects 
created by functions of this package into your preferred data object should be 
fairly easy. Note `xts` is a dependency listed under the "Suggests" field of 
`data.table` DESCRIPTION file.

For more information on `xts` objects, go [here](https://cran.r-project.org/web/packages/xts/index.html)
and [here.](http://rstudio-pubs-static.s3.amazonaws.com/288218_117e183e74964557a5da4fc5902fc671.html)


## yth_glm

The `yth_glm` function primarily exists to model the output for the `yth_filter`. 
However, one can use this wrapper around `glm` independently. On that note,
the function API allows you to use the `...` to pass any additional arguments
to `glm`.

The `yth_filter` returns an object of class `glm`, so one can use all 
generic methods associated with objects of that class. Here is an example of 
passing the results of a `yth_glm` model to the `plot` function, which outputs
the standard plot diagnostics associated with the method.

```{r, warning=FALSE, message=FALSE}
library(neverhpfilter)

data(GDPC1)

log_RGDP <- 100*log(GDPC1)

gdp_model <- yth_glm(log_RGDP["1960/"], h = 8, p = 4)

plot(gdp_model)
```


## yth_filtered

This is the main function of the package and both accepts, and outputs an `xts` 
object. The resulting output contains various series discussed in Hamilton (2017). 
These are a user defined combination of the original, trend, cycle, and random 
walk series. See documentation and the original paper for further details.

```{r, warning=FALSE, message=FALSE}
gdp_filtered <- yth_filter(log_RGDP, h = 8, p = 4)

tail(gdp_filtered, 16)

class(gdp_filtered)
```

The output is an `xts` object, it inherits all generic methods and capabilities
associated with `xts`. For example, one can conveniently produce clean times 
series graphics with `plot.xts`. 

Not the use of `xts::addPanel` function, which adds a panel plot of the 
`cycle` output of the `yth_filter` to the original graph.

```{r, warning = FALSE}
plot(log_RGDP, grid.col = "white", col = "blue", legend.loc = "topleft", main = "100 x Log of Real GDP (GDPC1)")
addPanel(yth_filter, output=c("cycle"), type="h", on=NA, col="darkred" )
```


## Choices for h and p

In the paper that inspired this package, Hamilton converts the `PAYEMS` series 
into quarterly periodicity and then uses filters on the transformed time series. 
With the `yth_filter` function, one can choose to retain the monthly series and 
adjust the `h` and `p` parameters accordingly. The default parameters of `h = 8` 
and `p = 4` assume times series data of a quarterly periodicity. For time series
of monthly periodicity, one can retain the same look-ahead and lag periods with 
`h = 24` and `p = 12`. See the `yth_filter` documentation for more details.

```{r, warning = FALSE}
Employment_log <- 100*log(PAYEMS["1950/"])

employment_cycle <- yth_filter(Employment_log, h = 24, p = 12, output = "cycle")

plot(employment_cycle, grid.col = "white", type = "h", up.col = "darkgreen", dn.col = "darkred", 
     main = "Log of Employment cycle")
```

In addition to adjust for various periodicities, one may wish to consider 
longer term cycles by extending `h`. Below are examples of moving the look-ahead
period defined by `h` from 8 quarters (2 years), to 20 quarters (5 years), and 
then 40 quarters (10 years). 

```{r}
gdp_5yr <- yth_filter(log_RGDP, h = 20, p = 4, output = c("x", "trend", "cycle"))

plot(gdp_5yr["1980/"][,1:2], grid.col = "white", legend.loc = "topleft", 
     main = "Log of Real GDP and 5-year trend", 
     panels = 'lines(gdp_5yr["1980/"][,3], type="h", on=NA)')

gdp_10yr <- yth_filter(log_RGDP, h = 40, p = 4, output = c("x", "trend", "cycle"))

plot(gdp_10yr["1980/"][,1:2], grid.col = "white", legend.loc = "topleft", 
     main = "Log of Real GDP and 10-year trend",
     panels = 'lines(gdp_10yr["1980/"][,3], type="h", on=NA)')

```

## Conclusion 

These functions give you the option to, in the words of Hamilton, never use the 
Hodrick-Prescott filter.They identify a more stable trend component that doesn't 
have the estimation issues associated with the `head` and `tail` portions of 
components generated by the HP filter. 
