---
title: "Economic Time Series Filtering: An alternative approach with the neverhpfilter package"
author: "Justin M. Shea"
date: "June 1st, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

- about me, contact information.

## Motivation

James Hamilton's working paper, 
**WHY YOU SHOULD NEVER USE THE HODRICK-PRESCOTT FILTER** [(2017)<doi:10.3386/w23429>](https://www.nber.org/papers/w23429) summarizes
the problem with the popular filter in three points:

- (1) The HP filter produces series with spurious dynamic relations that have no 
basis in the underlying data-generating process.

- (2) Filtered values at the end of the sample are very different from those in the middle, and are also characterized by spurious dynamics.  

- (3) A statistical formalization of the problem typically produces values for the smoothing parameter vastly at odds with common practice, e.g., a value for $\lambda$ far below **1600** for quarterly data.  


## If you gave someone HP-Filtered data feeling like this...

```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/top-gun.jpg")
knitr::include_graphics(image)
```



## ...In reality, its more like this:


```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/danger-zone.jpg")
knitr::include_graphics(image)
```


## The 4th point of Hamilton's abstract presents a solution

- (4) There's a better alternative. A regression of the variable at date **$t + h$** on the four most recent values (for quarterly data) as of date $t$ offers a robust approach to detrending that achieves all the objectives sought by users of the HP filter with none of its drawbacks.


$$y_{t+8} = \beta_0 + \beta_1 y_t + \beta_2 y_{t-1} +\beta_3 y_{t-2} + \beta_4 y_{t-3} + v_{t+8}$$

Which can be rewritten as:

$$y_{t} = \beta_0 + \beta_1 y_{t-8} + \beta_2 y_{t-9} + \beta_3 y_{t-10} + \beta_4 y_{t-11} + v_{t}$$

## Do any of Hamilton's peers agree? From the cover page:


_I thank Daniel Leff for outstanding research assistance on this project and **Frank Diebold, Robert King, James Morley**, and anonymous referees for helpful comments on an earlier draft of this paper._


## Implementing Hamilton's alternative: neverhpfilter package

The package consists of 2 core functions which are:

`yth_glm`: fits a generalized linear model object of class **`glm`**.

```{r, echo=TRUE,eval=FALSE}
yth_glm(x, h = 8, p = 4, ...)
```

`yth_filter`: returns an **`xts`** object containing user defined combinations of the original, trend, cycle, and random walk series.

```{r, echo=TRUE, eval=FALSE}
yth_filter(x, h = 8, p = 4, 
           output = c("x", "trend", "cycle", "random"), 
           ...)
```

 In addition the package comes with **14** documented data sets used to reproduce the results of Hamilton(2017).
 
## Hamilton's alternative: Quick Example

```{r, echo=TRUE}
library(neverhpfilter)
gdp_filtered <- yth_filter(100*log(GDPC1), h = 8, p = 4)
tail(gdp_filtered, 8)
class(gdp_filtered)
```


## Got Dependencies?


```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/dirk-josh.PNG")
knitr::include_graphics(image)
```
```{r, echo=TRUE, tidy=TRUE}
tools::package_dependencies("neverhpfilter") 
```


## Why depend on xts?

Model functions accept and return xts objects of any periodicity.

```{r, echo=TRUE}
class(GDPC1)
xts::periodicity(GDPC1)
y <- yth_filter(GDPC1)
class(y)
xts::periodicity(y)
```

## What else is awesome about xts?
`plot(x, ... )` quickly produces nice graphs.

```{r}
main <- "Log of Real GDP (GDPC1) and trend"
plot(gdp_filtered["1980/"][,1:2], grid.col = "white", legend.loc = "topleft", main = main,  panels = 'lines(gdp_filtered["1980/"][,3], type="h", on=NA)')
```


## Reproducing Hamilton's solution

```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/reproducing.PNG")
knitr::include_graphics(image)
```


## James Hamilton is cool with open source R


```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/Hamilton-email.PNG")
knitr::include_graphics(image)
```

## And he did it!

```{r echo=FALSE, out.width='100%'}
image <- paste0(getwd(),"/images/Hamilton-web.PNG")
knitr::include_graphics(image)
```




## Example

Scatterplot of correlation of Asset returns vs Growth/Inflation percent. 

Then show another graph of asset returns vs cycle or trend of GDP.

## Model estimation function

For model estimation, I settled on `glm` because, model object!

```{r, echo=TRUE}
gdp_model <- yth_glm(100*log(GDPC1), h = 8, p = 4)
```
```{r}
knitr::kable(broom::tidy(gdp_model))
```







